---
title: "Examining background gene lists using simulation - comparison of different enrichment techniques"
author: "Mark Ziemann & Anusuiya Bora"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 7
    fig_height: 5
theme: cosmo
---

Source: https://github.com/markziemann/noise

```{r,libs}

RhpcBLASctl::blas_set_num_threads(1)

suppressPackageStartupMessages({
library("parallel")
library("edgeR")
library("DESeq2")
library("limma")
library("stringi")
library("kableExtra")
library("fgsea")
library("clusterProfiler")
source("simpw_func.R")
})

```

## Intro

TODO


xxx object slots

1. expression counts

2. ground truth up genes

3. ground truth down genes

4. ground truth up gene sets

5. ground truth down gene sets

6. DE result (DESeq2)

7. DE genes up observed

8. DE genes down observed

9. fora up gene sets 5% FDR

10. fora down gene sets 5% FDR

11. fora up gene sets 1% FDR

12. fora down gene sets 1% FDR

13. fgsea up gene sets 5% FDR

14. fgsea down gene sets 5% FDR

15. fgsea up gene sets 1% FDR

16. fgsea down gene sets 1% FDR

# Get count data

```{r,getcounts,warning=FALSE}

a <- countData()

```

# Generate gene sets

```{r,genesetgeneration}

gsets <- randomGeneSets(a,setsize=30,nsets=200)

```

# run simulations over a range of parameters

```{r,setparameters}

SIMS=1000
FRAC_DE=0.05
FC=0.5
N_REPS=5
DGE_FUNC="deseq2"
SUM_COUNT=1e7
VARIANCE=c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1.0)

mygrid <- expand.grid(FRAC_DE,FC,N_REPS,DGE_FUNC,SUM_COUNT,VARIANCE)
colnames(mygrid) <- c("FRAC_DE","FC","N_REPS","DGE_FUNC","SUM_COUNT","VARIANCE")

mygrid

```

Now run the analysis.

```{r,sims}

if ( file.exists("simpw.Rds") ) {
  message("res.Rds found on disk, skipping")
} else {
res <- lapply(1:nrow(mygrid), function(i) {
  FRAC_DE=mygrid[i,"FRAC_DE"]
  FC=mygrid[i,"FC"]
  N_REPS=mygrid[i,"N_REPS"]
  DGE_FUNC=as.character(mygrid[i,"DGE_FUNC"])
  SUM_COUNT=mygrid[i,"SUM_COUNT"]
  VARIANCE=mygrid[i,"VARIANCE"]
  x <- agg_dge(a,N_REPS,SUM_COUNT,VARIANCE,FRAC_DE,FC,SIMS,DGE_FUNC,gsets,cores=16)
  as.data.frame(do.call(rbind, x))
})

res <- do.call(rbind,res)

saveRDS(res,file="simpw.Rds")
}

```

Now show the results.

```{r,results}

res %>% kbl(caption="simulation_results") %>% kable_paper("hover", full_width = F)

```

## Plot


```{r,plot1,fig.width=10,fig.height=6}

fo5 <- subset(res,PWAY_FUNC == "fora5")
fo1 <- subset(res,PWAY_FUNC == "fora1")
fg5 <- subset(res,PWAY_FUNC == "fgsea5")
fg1 <- subset(res,PWAY_FUNC == "fgsea1")

fo5 %>% kbl(caption="fora 5% FDR") %>% kable_paper("hover", full_width = F)
fg5 %>% kbl(caption="fgsea 5% FDR") %>% kable_paper("hover", full_width = F)
fo1 %>% kbl(caption="fora 1% FDR") %>% kable_paper("hover", full_width = F)
fg1 %>% kbl(caption="fgsea 1% FDR") %>% kable_paper("hover", full_width = F)

par(mfrow=c(1,3))
par(mar=c(c(5.1, 5.1, 2.1, 2.1) ))

plot(fo5$VARIANCE,fo5$p,ylim=c(0,1),type="b",pch=19,xlab="variance added",ylab="index",main="precision",col="red",lwd=3)
points(fg5$VARIANCE,fg5$p,type="b",pch=19,col="blue",lwd=3)
points(fo1$VARIANCE,fo1$p,type="b",pch=19,col="pink",lwd=3)
points(fg1$VARIANCE,fg1$p,type="b",pch=19,col="lightblue",lwd=3)
grid()

legend("bottomleft", inset=.02, title="tool",
   c("fora 5%","fgsea 5%", "fora 1%", "fgsea 1%"),
   col=c("red","blue","pink","lightblue"),horiz=FALSE, cex=1.0, pch=19,lwd=2)

plot(fo5$VARIANCE,fo5$r,ylim=c(0,1),type="b",pch=19,xlab="variance added",ylab="index",main="recall",col="red",lwd=3)
points(fg5$VARIANCE,fg5$r,type="b",pch=19,col="blue",lwd=3)
points(fo1$VARIANCE,fo1$r,type="b",pch=19,col="pink",lwd=3)
points(fg1$VARIANCE,fg1$r,type="b",pch=19,col="lightblue",lwd=3)
grid()

plot(fo5$VARIANCE,fo5$f,ylim=c(0,1),type="b",pch=19,xlab="variance added",ylab="index",main="F1",col="red",lwd=3)
points(fg5$VARIANCE,fg5$f,type="b",pch=19,col="blue",lwd=3)
points(fo1$VARIANCE,fo1$f,type="b",pch=19,col="pink",lwd=3)
points(fg1$VARIANCE,fg1$f,type="b",pch=19,col="lightblue",lwd=3)
grid()

##########################
png("sim.png", width=7,height=5,units="in",res=150,pointsize=12)
par(mar=c(c(5.1, 5.1, 2.1, 2.1) ))
par(mfrow=c(1,3))

plot(fo5$VARIANCE,fo5$p,ylim=c(0,1),type="b",pch=19,xlab="variance added",ylab="index",main="precision",col="red",lwd=3)
points(fg5$VARIANCE,fg5$p,type="b",pch=19,col="blue",lwd=3)
points(fo1$VARIANCE,fo1$p,type="b",pch=19,col="pink",lwd=3)
points(fg1$VARIANCE,fg1$p,type="b",pch=19,col="lightblue",lwd=3)
grid()

legend("bottomleft", inset=.02, title="tool",
   c("fora 5%","fgsea 5%", "fora 1%", "fgsea 1%"),
   col=c("red","blue","pink","lightblue"),horiz=FALSE, cex=1.0, pch=19,lwd=2)

plot(fo5$VARIANCE,fo5$r,ylim=c(0,1),type="b",pch=19,xlab="variance added",ylab="index",main="recall",col="red",lwd=3)
points(fg5$VARIANCE,fg5$r,type="b",pch=19,col="blue",lwd=3)
points(fo1$VARIANCE,fo1$r,type="b",pch=19,col="pink",lwd=3)
points(fg1$VARIANCE,fg1$r,type="b",pch=19,col="lightblue",lwd=3)
grid()

plot(fo5$VARIANCE,fo5$f,ylim=c(0,1),type="b",pch=19,xlab="variance added",ylab="index",main="F1",col="red",lwd=3)
points(fg5$VARIANCE,fg5$f,type="b",pch=19,col="blue",lwd=3)
points(fo1$VARIANCE,fo1$f,type="b",pch=19,col="pink",lwd=3)
points(fg1$VARIANCE,fg1$f,type="b",pch=19,col="lightblue",lwd=3)
grid()

dev.off()

#################################
pdf("sim.pdf", width=7,height=5,pointsize=12)
par(mar=c(c(5.1, 5.1, 2.1, 2.1) ))
par(mfrow=c(1,3))

plot(fo5$VARIANCE,fo5$p,ylim=c(0,1),type="b",pch=19,xlab="variance added",ylab="index",main="precision",col="red",lwd=3)
points(fg5$VARIANCE,fg5$p,type="b",pch=19,col="blue",lwd=3)
points(fo1$VARIANCE,fo1$p,type="b",pch=19,col="pink",lwd=3)
points(fg1$VARIANCE,fg1$p,type="b",pch=19,col="lightblue",lwd=3)
grid()
legend("bottomleft", inset=.02, title="tool",
   c("fora 5%","fgsea 5%", "fora 1%", "fgsea 1%"),
   col=c("red","blue","pink","lightblue"),horiz=FALSE, cex=1.0, pch=19,lwd=2)

plot(fo5$VARIANCE,fo5$r,ylim=c(0,1),type="b",pch=19,xlab="variance added",ylab="index",main="recall",col="red",lwd=3)
points(fg5$VARIANCE,fg5$r,type="b",pch=19,col="blue",lwd=3)
points(fo1$VARIANCE,fo1$r,type="b",pch=19,col="pink",lwd=3)
points(fg1$VARIANCE,fg1$r,type="b",pch=19,col="lightblue",lwd=3)
grid()

plot(fo5$VARIANCE,fo5$f,ylim=c(0,1),type="b",pch=19,xlab="variance added",ylab="index",main="F1",col="red",lwd=3)
points(fg5$VARIANCE,fg5$f,type="b",pch=19,col="blue",lwd=3)
points(fo1$VARIANCE,fo1$f,type="b",pch=19,col="pink",lwd=3)
points(fg1$VARIANCE,fg1$f,type="b",pch=19,col="lightblue",lwd=3)
grid()

dev.off()

```

Barplot is simpler.

```{r,bars}

prec <- c("fora 5%"=mean(fo5$p),"fgsea 5%"=mean(fg5$p),"fora 1%"=mean(fo1$p),"fgsea 1%"=mean(fg1$p))
rec <- c("fora 5%"=mean(fo5$r),"fgsea 5%"=mean(fg5$r),"fora 1%"=mean(fo1$r),"fgsea 1%"=mean(fg1$r))
f1 <- sapply(1:length(prec), function(i) { 2/(1/rec[i]+1/prec[i]) })

par(mfrow=c(1,3))
par(mar=c(c(9.1, 3.5, 2.1, 1.1) ))
barplot(prec,ylim=c(0,1.1),las=2,main="precision",ylab="")
text((1:6*1.18)-0.45,prec+0.02,labels=signif(prec,3))
barplot(rec,ylim=c(0,1.1),las=2,main="recall",ylab="")
text((1:6*1.18)-0.45,rec+0.02,labels=signif(rec,3))
barplot(f1,ylim=c(0,1.1),las=2,main="F1",ylab="")
text((1:6*1.18)-0.45,f1+0.02,labels=signif(f1,3))

png("simbars.png", width=7,height=5,units="in",res=150,pointsize=12)
par(mar=c(c(9.1, 3.5, 2.1, 1.1) ))
par(mfrow=c(1,3))
barplot(prec,ylim=c(0,1.1),las=2,main="precision",ylab="")
text((1:6*1.18)-0.45,prec+0.02,labels=signif(prec,3))
barplot(rec,ylim=c(0,1.1),las=2,main="recall",ylab="")
text((1:6*1.18)-0.45,rec+0.02,labels=signif(rec,3))
barplot(f1,ylim=c(0,1.1),las=2,main="F1",ylab="")
text((1:6*1.18)-0.45,f1+0.02,labels=signif(f1,3))
dev.off()

#################################
pdf("simbars.pdf", width=7,height=5,pointsize=12)
par(mar=c(c(9.1, 3.5, 2.1, 1.1) ))
par(mfrow=c(1,3))
barplot(prec,ylim=c(0,1.1),las=2,main="precision",ylab="")
text((1:6*1.18)-0.45,prec+0.02,labels=signif(prec,3))
barplot(rec,ylim=c(0,1.1),las=2,main="recall",ylab="")
text((1:6*1.18)-0.45,rec+0.02,labels=signif(rec,3))
barplot(f1,ylim=c(0,1.1),las=2,main="F1",ylab="")
text((1:6*1.18)-0.45,f1+0.02,labels=signif(f1,3))
dev.off()

```

## Session information

```{r,session}

sessionInfo()

```
